---
title: "MovieLens Project"
author: "Ng Da Xuan"
date: "5/31/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction of dataset & Overview of the goal of current Project 

The dataset used in this project came from the work of Harper & Konstan (2015). The dataset set is collected from a total of 71567 users, consists of 10000054 ratings, 95580 tags, and 10681 movies. The goal of the current project is to develop a Machine Learning model for the dataset that minimises the RMSE.

The key steps taken to reach the goal were: 1)cleaning the data into a tidy format, 2) exploring the data and understanding the data structure, & 3) testing out different modelling approach to find the best predictive model (ie., lowest RMSE).

The following are the libraries used in exploring the data. 

```{r libraries, warning = FALSE, message=FALSE}
library(matrixStats)
library(tidyverse)
library(caret)
library(lubridate)
library(knitr)
library(rmarkdown)
library(data.table)
library(stringr)
```

## Methods & Analysis

### Step 1: Cleaning the data into a tidy format

* Download data
```{r data source}
dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

```

* Organising dataset into tidy format

```{r organise data, tidy = TRUE}
ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))
movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")
movies <- as.data.frame(movies) %>%mutate(movieId =as.numeric(levels(movieId))[movieId],
                                          title = as.character(title),
                                          genres = as.character(genres))
movielens <- left_join(ratings, movies, by = "movieId")
```

* Creating a validation dataset and edx dataset

```{r create dataset, tidy = TRUE, warning=FALSE,message=FALSE}
## Validation set will be 10% of MovieLens data
set.seed(1, sample.kind="Rounding")
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

## Make sure userId and movieId in validation set are also in edx set
validation <- temp %>% 
  semi_join(edx, by = "movieId") %>%
  semi_join(edx, by = "userId")

## Add rows removed from validation set back into edx set
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)
rm(dl, ratings, movies, test_index, temp, movielens, removed)
```

### Step 2: Exploring the edx dataset and understanding the data structure

* Exploring the edx dataset

```{r}
## Exploring the datatype and type of variables in the dataset.
str(edx)
head(edx)

## How many rows and columns are there in the edx dataset?
dim(edx)

## How many zeros were given as ratings in the edx dataset?
edx %>% filter(rating == 0) %>% tally()

## How many threes were given as ratings in the edx dataset?
edx %>% filter(rating == 3) %>% tally()

## How many different movies are in the edx dataset?
n_distinct(edx$movieId)

## How many different users are in the edx dataset?
n_distinct(edx$userId)

## How many movie ratings are in each of the following genres in the edx dataset?
# str_detect
genres = c("Drama", "Comedy", "Thriller", "Romance", "Action", "Sci-Fi", "Adventure", "Children", "Fantasy")
sapply(genres, function(g) {
  sum(str_detect(edx$genres, g))
})

## Which movie has the greatest number of ratings?
edx %>% group_by(movieId, title) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```

```{r, results="hold", message=FALSE}
## What are the five most given ratings in order from most to least?
edx %>% group_by(rating) %>% summarize(count = n()) %>% top_n(5) %>%
  arrange(desc(count))

edx %>%
  group_by(rating) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = rating, y = count)) +
  geom_line()
```


### Step 3: Testing out different modelling approach to find the best predictive model



You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Results

## Conclusion

## References

F. Maxwell Harper and Joseph A. Konstan. 2015. The MovieLens Datasets: History and Context. ACM Transactions on Interactive Intelligent Systems (TiiS) 5, 4, Article 19 (December 2015), 19 pages. DOI=http://dx.doi.org/10.1145/2827872
